---
title: "PRIDE Archive Anual Trends - Exploratory Data Analysis"
author: "Jose A. Dianes"
date: "20 March 2015"
output:
  html_document:
    keep_md: yes
    theme: cerulean
    toc: yes
---

## Requisities  

These are some of the libraries and configurations we will use in the 
following analysis.  

```{r, message=FALSE}
require(devtools)
install_github("PRIDE-R/prideR")
library(prideR)
require(ggplot2)
theme_set(theme_linedraw())
require(dplyr)
```

## A draft of the pipeline  

The two first parts are carried out in this document. The data visualisation part
will be generated and published separatelly and referenced from here.  

#### Data acquisition

We will use the [prideR package](http://pride-r.github.io/prideR) to retrieve 
data from all the submitted experiments up to date.  

#### Exploratory Data Analysis  

By using `R`, first we will prepare and explore data to be mined. Then we will
compute different summary statistics and generate some plots that will give us
some insight into the questions we want to answer in our visualisation.  

#### Data Visualisation  

For this data viz we will use [Tableau](http://www.tableau.com). We will generate some worksheets for the 
key charts, and then we will put them together in an *author-driven* linear
narrative that will go over time answering questions about MS2 proteomics 
research trends.  

## Data acquisition  

Using [prideR](http://pride-r.github.io/prideR), lets retrieve all the public 
datasets. We will use a large number as page size so we retrieve all of them.    

```{r}
all_projects <- search.list.ProjectSummary("",0,20000)
num_projects <- length(project.list(all_projects))
```

This time we get `r num_projects` projects in total. Lets convert the result 
object into a data frame, more suitable to work with the different R libraries.  

```{r}
all_projects.df <- as.data.frame(all_projects)
names(all_projects.df)
```

Let's have a look at the publication dates.  

```{r}
str(all_projects.df$publication.date)
```

We see they are already in date format (`POSIXct`) and contain month and day. But we are just
interested in years in our study of trends, so let's extract the year in a separate
variable in the data frame.  

```{r}
all_projects.df$publication.year <- as.factor(format(all_projects.df$publication.date, "%Y"))
table(all_projects.df$publication.year)
```

We have a lot of information in that summary table. But let's put it visually.  

```{r}
ggplot(data=all_projects.df, aes(x=publication.year)) +
    geom_bar()
```

Maybe we should exclude the present year from our trend analysis since the year
is not complete and will incorrectly influence the outcomes.  

```{r}
all_projects_up2014.df <- subset(all_projects.df, publication.year!="2015")
num_projects_up2014 <- nrow(all_projects2014.df)
```

There we are. We have a new dataframe with `r num_projects_up2014` projects 
up to 2014. I think we are ready to start doing some Exploratory Analysis.  

## Exploratory Data Analysis  

So we know how the number of publications has changed over the years. But we are
also interested in what people research during these years. From the names in our
data frame we saw that each project includes information regarding:  

* Species  
* Tissues  
* Modification names  
* Instrument names  

Let's first have a look at each of these variables as a whole, and later on we
can group them by years.  

```{r}
species_table <- sort(table(all_projects_up2014.df$species), decreasing=T)
species_table_frequent <- species_table[species_table>2]
num_species <- length(species_table)
num_species_frequent <- sum(species_table>2)
```

We see that we have up to `r num_species` different species. But most of them 
are rare species that appear just once or twice. Just `r num_species_frequent`
of them have been submitted more than twice. Let's plot them.    

```{r}

```



